Imminent range shifts in transboundary fish stocks challenge their management under climate change

### 2021-03-17 Scripts/Manuscript.Rmd
Ale. Thom. and Rash. Comments

---
title: "Early range shifts in transboundary fish stocks challenge their management under climate change"
author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern. ^3^Oeschger Centre for Climate Change Research, University of Bern, Switzerland. ^4^School of Environmental Sciences, University of Liverpool. ^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States'
output: bookdown::word_document2
reference_docx:  ../References/sci_template.docx
  # pdf_document: default
bibliography: ../References/ToE_Ref.bib
csl: ../References/science.csl
---

@@

-   corresponding author: [j.palacios\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email}

@@

Climate change is shifting the distribution of transboundary fish stocks between neighbouring countries. The time-frame of such shifts determine the extent of climate challenges to international fisheries governance. Here, using global climate-fisheries simulation modelling, we show that, by 2036, 68% of the world’s transboundary stocks are projected to shift their distributions significantly across international boundaries. By 2100, such shifts will occur in all countries under a ‘no mitigation’ scenario. Hotspots where transboundary stock shifts are earlier and larger coincide with countries that are highly dependent on fisheries for income and livelihood. The early occurrences of transboundary shifts highlight the urgency to re-formulate international fisheries policies to avoid disputes in fish stock sharing. Such bi-/multi-lateral discussions should start now as lead-time for policy changes is long.

@@

Over the last century, human activities have altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters, [@IPCC:2019tn]. Marine species' distributions reflect species' preferences for discrete environmental conditions [@Hutchinson:1957nt; @Nelson:2016kk]. As a result of climate change, many marine species have shifted their distributions towards higher latitude, deeper water or following local temperature gradients to track and stay within their optimal environmental 'niche'[@Poloczanska:2016kk; @Barange:2014kr]. The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt; @Barange:2014kr], impacting fisheries production and economies [@Sumaila:2019ew], and compromising our capacity to reach international sustainability goals within the 2030 Agenda such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu]. The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management [@Cheung:2018dh], including for species that cross international borders, i.e., shared stocks [@Pinsky:2018cb].

The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by Coastal States [@UnitedNations:1986tl]. As defined by the United Nations' Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (*iv*) discrete stocks that are only present on the high seas [@Munro:2004th]. This study focuses on transboundary stocks exploited by fisheries operating within EEZs. Countries are responsible for the management of stocks within their EEZs, under UNCLOS, and are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl]. A recent study estimates that, globally, there are 633 transboundary fish species representing 67% of identified exploited taxa. These species yielded an annual average of 48 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].

The effectiveness of fisheries management for transboundary species, and thus reaching the 2030 Agenda [@Pecl:2017hu], is challenged by shifts in stocks' distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id]. In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw]. While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017iua; @Spijkers:2017ij; @Crespo:2020cs]. Such patterns are expected to be exacerbated by continuing climate change [@Pinsky:2018cb; @Sumaila:2011bq]. The importance of transboundary stocks for both catch and fisheries revenue [@PalaciosAbrantes:2020tv] highlights their critical role in achieving the SDGs. Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important to developing climate resilient international ocean governance and achieving the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].

Here, we estimate the year in which 663 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variation and the intensity of said shift.  We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas). We employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks' distribution across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (RCP8.5; Methods). We find that the shared distribution of many transboundary stocks has already shifted beyond historical natural variability. Moreover, by 2030, the shared catch proportion of such stocks will exceed historical levels in most neighboring EEZs, posing a threat to the sustainability of associated fisheries and the resilience of international treaties that manage their shared exploitation.

@@

We calculated the time of emergence of the transboundary index to identify shifts in the shared distribution of transboundary stocks from 2006 to 2100 under a high emission climate change scenario. Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability as early as 2006 and up to 2100 (Fig. \@ref(fig:toeFigure1)A). This corresponds to 68% of the studied stocks (Fig. \@ref(fig:toeFigure1)C). On average, the shift in shared distribution of these stocks in all EEZs analyzed is projected to be 2036 $\pm$ 28 years. Furthermore, 96% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with around 80% presenting an early distributional shift across stocks (Fig. \@ref(fig:toeFigure1)C).

```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary stock shared distribution. Dashed lines represent the year by which countries have committed to reach the full implementation of the 2030 Agenda, as well as an early and late change in range shift, respectively."}

@@

The median year in which transboundary stocks experience a range shift varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Figure \@ref(fig:toeFigure2)A). Overall, most tropical EEZs will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia being significantly earlier (*p* < 0.05; see Table \@ref(tab:toeTableS2) for test statistics) than almost all other regions (Fig. \@ref(fig:toeFigure2)B). In contrast, EEZs located in temperate regions like northern Europe and eastern Asia are projected to experience significantly later range shifts (Table \@ref(tab:toeTableS2)). Stocks in some countries, like Brazil or New Zealand, are not projected to experience range shifts beyond natural variability (e.g., the transboundary index of a stock did not emerge from historical variability under a 95% confidence interval) (Fig. \@ref(fig:toeFigure2)A). On average, shifting transboundary stocks represented 23% $\pm$ 22% of yearly fishing revenue from fisheries targeting such stocks within global EEZs between 2004 and 2010 (Fig. \@ref(fig:toeFigure2)A). However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands). Moreover, in some EEZs, while few transboundary stocks are projected to shift above historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).

```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks. A) Land polygon shows the contribution of shifting stocks to the country’s total fishing revenue from transboundary stocks. Exclusive Economic Zone polygons represent average range shifts per EEZ. EEZs with no time of emergence between 2006 and 2100 are represented in aqua color. B) Year of shared distribution shifts by region according to United Nations sub-regions. Points represent the median year of range shift. Horizontal dashed line represents the 2030 Agenda. N = North, S = South, W = West and E = East. Ltn. Ame. and the Car. = Latin America and the Caribbean. Aus and New Z. = Australia and New Zealand"}

@@

We estimated the current (2005 and 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs. Within five years of the UN 2030 Agenda, 979 transboundary stocks responsible for 5.3 billion USD in fishing revenue are projected to shift within the EEZs of 76 counties (Fig. \@ref(fig:toeFigure3)); with 4.2 billion USDs accounted by 321 stocks shifting in just eleven countries (Fig. \@ref(fig:toeFigure3)). Most of the remaining countries in the top $75^{th}$ revenue percentile will see early shifts (e.g., before 2050) with the exception of Japan, Korea, Peru and Russia, who are projected to have later shifts (e.g., after 2050). In terms of the number of transboundary stocks shifting from historical variability, Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66), emerge as the top 5 countries (Fig. \@ref(fig:toeFigure3)).

```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks per fishing entity. Countries are coded according to different colored symbols representative of 5 distinct regions. Horizontal line separating the top 75th percentile of annual fishing revenue. Showing country names for those within the top 75th percentile projected to  have an average shift within 5 years of the 2030 Agenda."}

@@

We compared the year in which transboundary stocks experience a range shift by species’ preferred habitat association (Fig. \@ref(fig:toeFigure4)). We found significant differences in the range shift year relative to habitat association (Kruskal-Wallis, $X^2$ = 203.85, DF = 93, *p* < 0.001; See table S3 for full statistics \@ref(tab:TabelS3)). The average range shift of bathy-pelagic species (e.g., Blue whiting *Micromesistius poutassou*), pelagic-oceanic (e.g., tunas) and reef species (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to happen before 2030 while bathydemersal species (e.g., Banguela hake, *Merluccius polli*) are expected to experience the latest shift (2040).

```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association. Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1). The category “Other” consists of species that have no “habitat preference” classification. The number of stocks included in this analysis for each habitat type is noted in parentheses. Dashed whiskers represent the 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Solid line represents the median time of emergence per habitat preference."}

img <- readPNG("../Figures/Fig4.png")
grid.raster(img)

<!-- ![toeFigure 4. Comparison of shifts in the shared distribution of transboundary stocks by species’ habitat association. Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Table S1). The category “Other” consists of species that have no “habitat preference” classification. The number of stocks included in this analysis for each habitat type is noted in parentheses. Dashed whiskers represent the 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Solid line represents the median time of emergence per habitat preference.](../Figures/Fig5.png) -->

@@

We identified transboundary stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point by the early (2020-2040) and mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) under a high emission climate change scenario. Findings show that by the early $21^{st}$ century, the global average catch proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 59% $\pm$ 17% (Fig. \@ref(fig:toeFigure5)A). By the early $21^{st}$, 85% (n = 239) of the world EEZs would have experienced such changes. Changes in stock share ratio are expected to increase slightly towards the mid 21st century at both the stock and EEZ level (Fig. \@ref(fig:toeFigure5)A and \@ref(fig:toeFigureS5)).

@@

Our results underscore regional "hotspots" of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figures \@ref(fig:toeFigure2)A and \@ref(fig:toeFigure5)A). Such regions, like the Caribbean, are characterized by high levels of warming relative to natural internal variability [@IPCC:2019tn; @Hawkins:2012hc] and high vulnerability of species to warming waters [@IPCC:2014bz]. Moreover, they represent relatively small EEZs that border multiple countries (e.g., The Caribbean). Game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Gronbaek:2020ei], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting transboundary stocks.

@@

The early projected range shifts of the large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy]. Specifically, Frölicher et al., 2016 estimated the time of emergence of dissolved oxygen ($O_2$ between 100-600m), natural primary production (NPP), and sea surface temperature (SST) relative to historical variability (1986-2005). By 2030, the SST signal will have emerged from historical variability for 70% of the global ocean under RCP 8.5 [@Frolicher:2016dg]. Early emergence of SST will occur mainly in the tropics while the Southern Ocean and the North Atlantic are not expected to experience an emergence on the SST signal before 2100 [@Reygondeau:2020jh; @Frolicher:2016dg]. For NPP and $O_2$, the signal is projected to emerge in 23% and 10% of the ocean by the end of the $21^st$ century, respectively under RCP 8.5 [@Frolicher:2016dg]. Results from multiple marine ecosystem models show that SST and NPP are the main drivers of changes in species distributions across oceanic basins [@BryndumBuchholz:2019kh; @Lotze:2019ce]. Specifically for the DBEM, SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results [@Cheung:2016tt]. The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz], while also explaining the later, and sometimes non-existent, shift at higher latitudes (Fig. \@ref(fig:toeFigure2)B).

Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and  *ii*) model structure [@Frolicher:2016dg]. First, our analysis is based on a high emission scenario (RCP 8.5) representing an extreme case of shifts in the shared distribution of transboundary species. Thus, any climate change mitigation pathway (e.g., RCP 2.6) could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg]. Second, a substantial source of uncertainty is related to model selection for both climate change and fisheries distribution. This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to the emergence of environmental drivers is larger but that to scenarios is lower [@Frolicher:2016dg]. While such uncertainty could potentially be reduced by incorporating  ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] and they ignore other sources of structural uncertainty based on the parameterisation of processes. As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties, and will likely continue to do so in the coming decade. This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf] and where internal variability (e.g., El Niño Southern Oscillation - ENSO- for Pacific nations) is also important . These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh]. Overall, changes in fish biomass with multiple upper trophic level models, including the DBEM, show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@BryndumBuchholz:2019kh]. Further research that reduces uncertainty in the response of the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could elucidate a broader range in structural uncertainty of fish and fisheries models [@Lotze:2019ce]. Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks. While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca]. Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie]. Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.

>>>>

---
title: "Imminent range shifts in transboundary fish stocks challenge their management under climate change"
author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas L. Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern. ^3^Oeschger Centre for Climate Change Research, University of Bern, Switzerland. ^4^School of Environmental Sciences, University of Liverpool. ^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States'
output: bookdown::word_document2
reference_docx:  ../References/sci_template.docx
  # pdf_document: default
bibliography: ../References/ToE_Ref.bib
csl: ../References/science.csl
---

@@

corresponding author: [j.palacios\@oceans.ubc.ca](mailto:j.palacios@oceans.ubc.ca){.email}

@@

Climate change is shifting the distribution of transboundary fish stocks between neighbouring countries. The time-frame of such shifts determine the extent of climate change challenges to international fisheries governance. Using global climate-fisheries model simulations, we show that, within the next 15 years, 68% of the world’s transboundary fish stocks are projected to shift their distributions across international boundaries. By 2100, such shifts will occur in all countries under a ‘no mitigation’ scenario. Hotspots of transboundary shifts occur in countries that are highly dependent on fisheries for income and livelihood. The imminent onset of transboundary shifts highlight the urgency to re-formulate international fisheries policies to avoid disputes over fish stock.  Bi- and multi-lateral actions discussions should start now as lead-time for policy changes is long.

@@

In the past century, human activities have altered the physical and biogeochemical conditions of the ocean, resulting in warmer, more acidic and less oxygenated waters [@IPCC:2019tn]. Marine species' distributions reflect species' preferences for discrete environmental conditions [@Hutchinson:1957nt].As a result of these past changes, many marine species have shifted their distributions towards higher latitudes, deeper water or to follow local temperature gradients to remain within their optimal environmental "niche" [@Poloczanska:2016kk; @Barange:2014kr]. The biogeography of marine species is projected to continue to shift as ocean conditions change through the 21st century [@Cheung:2010dt; @Barange:2014kr], impacting fisheries production and economies [@Sumaila:2019ew], and compromising our capacity to reach international sustainability goals within the 2030 Agenda such as Sustainable Development Goal 14 - life below water [@UnitedNations2018; @Singh:2017ds; @Pecl:2017hu]. The projected risks of impacts can be reduced by improving the effectiveness of current fisheries management [@Cheung:2018dh], including for species that cross international borders, which are known as ‘shared stocks’ [@Pinsky:2018cb].

The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by Coastal States [@UnitedNations:1986tl]. As defined by the United Nations' Food and Agriculture Organization, shared stocks can be classified into four non-exclusive categories: (*i*) transboundary stocks, which cross neighboring EEZs; (*ii*) straddling stocks that, in addition to neighboring EEZs, also visit the adjacent high seas; (*iii*) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and (*iv*) discrete stocks that are only present in the high seas [@Munro:2004th]. This study focuses on transboundary stocks exploited by fisheries operating within EEZs. Countries are responsible for the management of stocks within their EEZs, under UNCLOS, and are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl]. A recent study estimates that, globally, there are 633 transboundary fish species representing 67% of identified exploited taxa. These species yielded an annual average of 48 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 [@PalaciosAbrantes:2020tv].

The effectiveness of fisheries management for transboundary species, and thus reaching the 2030 Agenda [@Pecl:2017hu], is challenged by shifts in stocks' distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by; @Pinsky:2020jx; @Oremus:2020id]. In most cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the full distribution range of the stock [@FredstonHermann:2018kp], nor the effects of climate change on fish stocks and associated fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw]. While many transboundary stocks experience natural seasonal spatial variation (e.g., migration, reproductive cycle), misalignment between fisheries resources allocation and distributional shifts beyond such natural variation have previously resulted in unsustainable fishing levels and international disputes [@Miller:2013iv; @Song:2017iua; @Spijkers:2017ij; @Crespo:2020cs]. Such patterns are expected to be exacerbated by continuing climate change [@Pinsky:2018cb; @Sumaila:2011bq]. The importance of transboundary stocks for both catch and fisheries revenue [@PalaciosAbrantes:2020tv] highlights their critical role in achieving the  Sustainable Development Goals (SDG) under the United Nations 2030 Agenda. Understanding when climate change will affect the sharing dynamics of transboundary stocks and the intensity of the resulting impacts is important for developing climate resilient international ocean governance and achieving the 2030 Agenda [@Pinsky:2018cb; @Link:2010ei; @Sumaila:2020tp].

Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth System Model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario (RCP8.5; Methods). We then estimate the year in which the 663 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variation and the intensity of this shift (i.e., time of emergence; see Methods). We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (15) and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).

@@

Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability as early as 2006 and up to 2100 (Fig. \@ref(fig:toeFigure1)A). This corresponds to 68% of the studied stocks (Fig. \@ref(fig:toeFigure1)C).The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be 2036 (s.d. of all species, $\pm$ 28 years). Furthermore, 96% of the world’s EEZs will experience at least one transboundary stock shift between 2006 and 2100 with around 80% presenting a distributional shift across stocks by the time the UN 2030 Agenda is intended to be achieved (Fig. \@ref(fig:toeFigure1)C).

```{r toeFigure1, eval = T,  echo = F, fig.width = 9, fig.height=4, fig.pos= "H", fig.cap =  "Timeline of shifts in transboundary shared stock distribution. Dashed lines represent the year by which countries have committed to reach the full implementation of the 2030 Agenda, as well as a mid and late change in range shift, respectively."}

@@

The median year in which transboundary stocks experience a range shift varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 242.11, DF = 93, *p* < 0.001; Figure \@ref(fig:toeFigure2)A). Overall, most tropical EEZs will see earlier shifts, with the EEZs of Latin America, the Caribbean and Polynesia being significantly earlier (*p* < 0.05; see Table S\@ref(tab:toeTableS2) for test statistics) than almost all other regions (Fig. \@ref(fig:toeFigure2)B). In contrast, EEZs located in temperate regions like northern Europe and eastern Asia are projected to experience much later range shifts (Table \@ref(tab:toeTableS2)).Stocks in countries such as Brazil and New Zealand are not projected to experience range shifts beyond natural variability (e.g., the transboundary index of a stock did not change beyond historical variability) (Fig. \@ref(fig:toeFigure2)A). On average, shifting transboundary stocks represented 23% (s.d. $\pm$ 22%) of yearly fishing revenue from fisheries targeting such stocks within global EEZs between 2004 and 2010 (Fig. \@ref(fig:toeFigure2)A). However, large variation exists, with shifting stocks representing less than 1% of fishing revenue in some countries (e.g., Ireland) and over 90% in others (e.g., Marshall Islands). Moreover, in some EEZs, while few transboundary stocks are projected to shift beyond historical variability, they still represent a large proportion of revenue derived from fisheries within that EEZ (e.g., Peru).

```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks. A) Land polygon shows the contribution of shifting stocks to the country’s total fishing revenue from transboundary stocks. Exclusive Economic Zone polygons represent average range shifts within them. EEZs with no distributional shifts between 2006 and 2100 are represented in aqua color. B) Year of shared distribution shifts by region according to United Nations sub-regions. Points represent the median year of range shift. Horizontal dashed line represents the year 2030. N = North, S = South, W = West and E = East. Ltn. Ame. and the Car. = Latin America and the Caribbean. Aus and New Z. = Australia and New Zealand"}

@@

We estimated the current (2005 and 2010) contribution of shifting stocks to yearly fishing revenue generated from fisheries targeting transboundary stocks within global EEZs. Within five years of the UN 2030 Agenda, 979 transboundary stocks responsible for 5.3 billion USD in fishing revenue are projected to shift within the EEZs of 76 countries (Fig. \@ref(fig:toeFigure3)); with 4.2 billion USDs accounted by 321 stocks shifting in just eleven countries (Fig. \@ref(fig:toeFigure3)). Most of the remaining countries in the top $75^{th}$ revenue percentile will see early shifts (e.g., before 2050) with the exception of Japan, Korea, Peru and Russia, who are projected to have later shifts (e.g., after 2050). In terms of the number of transboundary stocks shifting from historical variability, Spain (n = 122), France (n = 85), China and Portugal (both n = 73), and Senegal (n = 66), emerge as the top 5 countries (Fig. \@ref(fig:toeFigure3)).

```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average range shift of transboundary stocks per country. Countries are coded according to different colored symbols representative of 5 distinct regions. Horizontal line separating the top 75th percentile of annual fishing revenue. Showing country names for those within the top 75th percentile projected to  have an average shift within 5 years of the 2030 Agenda."}

@@

@@

We identified transboundary stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point by the early (2020-2040) and mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005). Findings show that by the early $21^{st}$ century, the global average catch proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 59% $\pm$ 17% (Fig. \@ref(fig:toeFigure5)A). By the early $21^{st}$, 85% (n = 239) of the world EEZs would have experienced such changes. Changes in stock share ratio are expected to increase slightly towards the mid 21st century at both the stock and EEZ level (Fig. \@ref(fig:toeFigure5)A and \@ref(fig:toeFigureS5)).

@@

Our results underscore regional "hotspots" of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figures \@ref(fig:toeFigure2)A and \@ref(fig:toeFigure5)A). Such regions, like the Caribbean, are characterized by high levels of warming relative to natural internal variability [@IPCC:2019tn; @Hawkins:2012hc] and high vulnerability of species to warming waters [@IPCC:2014bz]. Moreover, they represent relatively small EEZs that border multiple countries (e.g., the Caribbean). Game theory predicts that the greater the number of negotiating parties, the harder it is for parties to reach an agreement [@Gronbaek:2020ei], thus, making it particularly challenging for countries in these regions to coordinate the management of shifting transboundary stocks.

@@

The early projected range shifts of a large number of transboundary stocks can be partially attributed to the parallel global emergence of several ocean variables from historical variability (e.g., SST which is projected to increase historical variability by 2030 in 70% of the global ocean under RCP 8.5 [@Rodgers:2015jq; @Frolicher:2016dg; @Mahlstein:2011fy]. Results from multiple marine ecosystem models show that SST and NPP are the main drivers of changes in species distributions across oceanic basins [@BryndumBuchholz:2019kh; @Lotze:2019ce]. Specifically for the DBEM, SST is the main environmental driver of biomass changes in polar, tropical and upwelling ecosystems while NPP drives temperate regions’ results [@Cheung:2016tt]. The combination of the SST emergence pattern and model characteristics could be partially responsible for the early distributional shift of transboundary stocks in the tropics where marine species live close to their thermal tolerance, making them highly vulnerable to warming waters [@IPCC:2014bz], while also explaining the later, and sometimes non-existent, shift at higher latitudes (Fig. \@ref(fig:toeFigure2)B).

Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and  *ii*) model structure [@Frolicher:2016dg]. First, our analysis is based on a high emission scenario (RCP 8.5) representing an extreme case of shifts in the shared distribution of transboundary species. Thus, any climate change mitigation pathway (e.g., RCP 2.6) could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg]. Second, a substantial source of uncertainty is related to model selection for both climate change and fisheries distribution. . This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to model selection, including the parameterization of poorly understood processes that regulate NPP changes, is larger than the uncertainty stemming from the climate change scenario [@Frolicher:2016dg]. While such uncertainty could potentially be reduced by incorporating ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] that are only just becoming available **(Deser et al. 2020)** and cannot address limitations of process level understanding. As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties, and will likely continue to do so in the coming decade. This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf] and where internal variability (e.g., El Niño Southern Oscillation - ENSO- for Pacific nations) is also important . These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh]. Overall, changes in fish biomass with multiple upper trophic level models, including the DBEM, show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@BryndumBuchholz:2019kh]. Further research that reduces uncertainty in the response of the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could lead to a larger structural uncertainty of fish and fisheries models [@Lotze:2019ce]. Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks. While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca]. Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie]. Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
